Do motif roles provide unique information about species risk of extinction?\\ \medskip Supplementary Information

### 2021-09-29 manuscript/supplement_motifs.tex
working on SI methods still

        These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
        We centered and scaled the predictors before fitting the LMMs.
        The models also included a random intercept for the interaction between species richness and connectance (equations.
        All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
        We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.

        The formula for these LMMs are:

        where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.

        We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
        To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
        These regressions were fit using the R~\citep{R} base function `lm'.
        We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance. 
        We performed similar analyses to test whether appearing in different positions in each motif (i.e., the motif role) is related to a species' persistence (\emph{Appendix SH}).

    \subsection{How does motif participation vary with species and network properties?}

        As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
        To establish this context for motif participation, we fit a linear mixed-effect model relating the frequency of participation in each motif to network size, connectance, and their interaction, as for network motif profiles (\emph{Appendix SG}).
        We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
        We performed similar analyses to test whether a species' motif role is related it in-degree or trophic lev,el (\emph{Appendix SI}).

>>>>

        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
        We centered and scaled the predictors before fitting the LMMs.
        All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
        We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.

        We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
        To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources:

        \begin{equation}
            \Psi_{ik} \approx \rho_{i} ,
            \label{mineq}
        \end{equation}

        where all symbols are as in equation~\ref{propreq}.
        These regressions were fit using the R~\citep{R} base function `lm'.
        We then visually examine the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.

    \subsection{How does motif participation vary with network properties?}

        As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
        To establish this context for motif participation, we first fit a PERMANOVA~\citep{Anderson2001} to test whether a species' overall motif participation role is related to network size, connectance, or the interaction between them.
        To avoid inflating similarity between species which do not appear in the same motifs (i.e., which have ``shared zeros''), we use Bray-Curtis dissimilarity between motif participation roles.
        Significance was calculated using 9999 permutations.
        The PERMANOVA was fit using the R~\citep{R} function `adonis' from the package \emph{vegan}~\citep{vegan}.

        To obtain a more detailed picture of the relationships between motif participation and global network structure, we fit a set of four linear models (one per motif):

        \begin{equation}
            \rho_{i} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i} ,
            \label{partic_SC}
        \end{equation}

        where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
        $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
        Predictors were not centered or scaled before fitting.
        All models were fit using the R~\citep{R} base function `lm'.

        We fit a similar set of eight linear mixed-effect models relating motif participation to in-degree or trophic level:

        \begin{equation}
            \rho_{i} \approx \delta_{i} + S_{i}C_{i} ,
            \label{partic_deg}
        \end{equation}

        \begin{equation}
            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
            \label{partic_deg}
        \end{equation}

        where $\rho_{i}$ is the frequency with which species $i$ participtes in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{i}C_{i}$ is a random effect of the combination of size and connectance for the network containing species $i$.
        These models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
